Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-09-25 prob.tex
prob.tex \subsection{Examples for Quantum Interval-valued Probability Measures} Updating...

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify this example is convex!}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined as follow. When the event is a 1-dimensional projector~$\proj{\psi}$,
it is also illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps})=\necess$,
where $\ket{0}$ and $\ket{\ps}$ is illustrated as the red and green
dotted vector, respectively.
\item Consider the states orthogonal to $\ket{0}$ and $\ket{\ps}$ respectively,
$\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. $\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}$ are points in the red and
green circles, and perpendicular to $\ket{0}$ and $\ket{\ps}$, respectively.
The dotted half of those states need special treatment, i.e., when
where $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we define
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely\textrm{ .}
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi})=\likely$. 
\end{enumerate}
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation of cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow. 
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

To verify $\bar{\mu}$ is a quantum interval-valued probability measure,
it is sufficient to check 
\begin{equation}
\begin{aligned}\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{2}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{3}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{1}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{3}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)
\end{aligned}
\label{eq:non-additive-vectors}
\end{equation}
for every orthonormal basis $\mathcal{B}=\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $.
We are going to enumerate all possible orthonormal bases to verify
the equations, and case 1 and 2 are illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{figure}
\includegraphics{\string"Three-dimensional quantum 4-value interval-valued probability measure\string".pdf}\caption{\label{fig:three-dimensional-4-value}This figure illustrates case
1 and 2 in $\mathbb{R}^{3}$ in example~\ref{ex:three-dimensional-4-value}.
The red and green dotted vector are $\ket{0}$ and $\ket{\ps}$ respectively.
All possible vectors of $\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\ps_{\theta,\gamma}^{\perp}}$
are drawn in the red and green circles, respectively. Within the circles,
a vector~$\ket{\psi}$ in dotted part means $\bar{\mu}(\proj{\psi})=\likely$;
otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is
a generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red
and green solid vectors are normalized $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively.}
\end{figure}

>>>>

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify this example is convex!}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined as follow, where the state vectors corresponding to 1-dimensional
projector are also illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{enumerate}
\item We have 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps}) & = & \necess\textrm{ ,}
\end{eqnarray*}
 where $\ket{0}$ and $\ket{\ps}$ is illustrated as the red and green
dotted vector, respectively.
\item Consider the states orthogonal to $\ket{0}$ and $\ket{\ps}$ respectively,
$\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$, as the
red and green circles in figure~\ref{fig:three-dimensional-4-value}.
The dotted half of those states need special treatment, i.e., when
where $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we define
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely\textrm{ .}
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi})=\likely$. 
\end{enumerate}
Assume there is a real-valued probability measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have is $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
A contradiction! Hence, we have $\mu(P)\notin\bar{\mu}(P)$ for all
quantum probability measure~$\mu$.

Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation of cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow:
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

Similar to example~\ref{ex:three-dimensional-three-value}, to verify
$\bar{\mu}$ is a quantum interval-valued probability measure, we
only need to verify equation~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
and it is sufficient to check 
\begin{equation}
\begin{aligned}\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{2}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{3}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{1}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{3}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)
\end{aligned}
\label{eq:non-additive-vectors}
\end{equation}
for every orthonormal basis $\mathcal{B}=\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $.
We are going to enumerate all possible orthonormal bases to verify
the equations, and case 1 and 2 are illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{figure}
\includegraphics{\string"Three-dimensional quantum 4-value interval-valued probability measure\string".pdf}\caption{\label{fig:three-dimensional-4-value}This figure illustrates case
1 and 2 in $\mathbb{R}^{3}$ in example~\ref{ex:three-dimensional-4-value}.
The red and green dotted vector are $\ket{0}$ and $\ket{\ps}$ respectively.
All possible vectors of $\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\ps_{\theta,\gamma}^{\perp}}$
are drawn in the red and green circles, respectively. Within the circles,
a vector~$\ket{\psi}$ in dotted part means $\bar{\mu}(\proj{\psi})=\likely$;
otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is
a generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red
and green solid vectors are normalized $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively.}
\end{figure}
